---
title: "Problem Set 2"
author: "Amy Tan"
format: pdf
editor: visual
---

## Github

Github link: <https://github.com/amyqtan/gov51-ps2>

## Loading Packages & Data

```{r}
library(tidyverse) 
gay <- read_csv("data/raw/gay.csv")
```

## Q1.1

```{r}
nrow(gay) #Number of total observations
gay %>% count(study)
gay %>% count(treatment)

```

There are 69592 total observations. The study takes two unique values. All treatment conditions in the treatment variable are No Contact, Recycling Script by Gay Canvasser, Recycling Script by Straight Canvasser, Same-Sex Marriage Script by Gay Canvasser, and Same-Sex Marriage Script by Straight Canvasser. (There are five total treatment conditions.)

## Q1.2

```{r}
gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  group_by(treatment) %>%
  summarize(mean(ssm), n())

```

For Study 1, Wave 1, treatment group "Recycling Script by Gay Canvasser" shows highest average support.

## Q1.3

```{r}
gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  group_by(treatment) %>%
  summarize(sd(ssm))
```

The standard deviations appear to be roughly similar across groups. This is important because similar standard deviations show that the randomization process was actually random and that pre-treatment, the different groups have similar baseline characteristics and values. This helps in the end, because when there are differences post-treatment, these differences can be attributed to the treatment itself and not pre-treatment differences or variations since th.

## Question 1.4

```{r}
study1wave1_mean <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  group_by(treatment) %>%
  summarize(mean_ssm = mean(ssm))

study1wave1_plot <- study1wave1_mean %>% 
  ggplot(mapping = aes(x=mean_ssm, y = treatment)) + geom_col(stat="identity", fill = "blue") + labs(title = "Mean Support for Gay Marriage by Group", y = "Treatment", x = "Mean Same Sex Marriage Attitude") 

study1wave1_plot
```

## Question 1.5

```{r}
mean_marriage_gay_w1 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(mean(ssm))

sd_marriage_gay_w1 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(sd(ssm))

mean_no_contact_w1 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  filter(treatment == "No Contact") %>%
  summarize(mean(ssm))

sd_no_contact_w1 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 1) %>%
  filter(treatment == "No Contact") %>%
  summarize(sd(ssm))

mean_marriage_gay
sd_marriage_gay
mean_no_contact
sd_no_contact

std_diff_means <- (mean_marriage_gay_w1 - mean_no_contact_w1)/sqrt(((sd_marriage_gay_w1)^2 + (sd_no_contact_w1)^2)/2)

std_diff_means
```

These groups look similar at baseline, since the standard difference of means is -0.010, which has an absolute value lower than 0.25. Since the absolute value of the standard difference in means is below 0.25, it suggests good balance. Moreover, it can be observed that the means and standard deviations for both groups are similar too, further supporting this claim.

## Question 1.6

```{r}
mean_marriage_gay_w2 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 2) %>%
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(mean(ssm))

sd_marriage_gay_w2 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 2) %>%
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(sd(ssm))

mean_no_contact_w2 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 2) %>%
  filter(treatment == "No Contact") %>%
  summarize(mean(ssm))

sd_no_contact_w2 <- gay %>% 
  filter(study == 1) %>% 
  filter(wave == 2) %>%
  filter(treatment == "No Contact") %>%
  summarize(sd(ssm))


std_diff_means <- (mean_marriage_gay_w2 - mean_no_contact_w2)/sqrt(((sd_marriage_gay_w2)^2 + (sd_no_contact_w2)^2)/2)


std_diff_means
```

## Question 2.1

```{r}
gay_r <- read_csv("data/raw/gayreshaped.csv") 
ccap <- read_csv("data/raw/ccap2012.csv")

nrow(gay_r)
colnames(gay_r)
nrow(ccap)
colnames(ccap)
```

For the gay reshaped dataset (gay_r), there are 11948 total observations and the column names are "study", "treatment", "therm1", "therm2", "therm3", and "therm4".

For the CCAP 2012 dataset (ccap), there are 43998 total observations and the column names are "...1", "caseid", and "gaytherm".

## Question 2.2

```{r}
comparison_table <- bind_rows(
  gay_r %>% 
  filter(study == 1) %>% 
  summarize(dataset = "Gay Replication",
            mean = mean(therm1), 
            median = median(therm1), 
            sd = sd(therm1)), 
  
  ccap %>% 
  na.omit(ccap$gaytherm) %>% 
  summarize(dataset = "CCAP 2012",
            mean = mean(gaytherm), 
            median = median(gaytherm), 
            sd = sd(gaytherm)))

comparison_table
```

writi

## Question 2.3

```{r}
library(ggplot2)

gay_r_ccap <- bind_rows(gay_r, ccap, .id = "source") #gay_r is 1, ccap is 2

gay_r_ccap <- gay_r_ccap %>%
  mutate(combined_therm = coalesce(therm1, gaytherm))

gay_r_ccap <- gay_r_ccap %>%
  mutate(datasource = case_when(source == 1 ~ "Gay Replication", source == 2 ~ "CCAP 2012"))

comparison_graph <- gay_r_ccap %>%
  ggplot(mapping = aes(x=combined_therm)) + geom_histogram(breaks = seq(0, 100, by = 5), color ="black", fill = "steelblue", binwidth = 5) + facet_wrap(~datasource) + 
    labs(x = "Thermometer Score towards Gay Men and Lesbians",
         y = "Number of Respondents",
         title = "Side-by-Side Comparison of Reshaped LaCour Data and CCAP 2012 Data") +
   theme_minimal() 
comparison_graph

ggsave("/Users/amytan/Desktop/gov51-ps2/output/commute_histogram.png", plot = comparison_graph)
```

## Question 2.4a

```{r}
gay_r_study1 <-gay_r %>% 
  filter(study == 1)

cor_therm_study1 <- cor(gay_r_study1$therm1, gay_r_study1$therm2, use = "complete.obs")
cor_therm_study1

gay_r_study2 <-gay_r %>% 
  filter(study == 2)

cor_therm_study2 <- cor(gay_r_study2$therm1, gay_r_study2$therm2, use = "complete.obs")
cor_therm_study2

```

## Question 2.4b

```{r}
gay_r_study1 <-gay_r %>% 
  filter(study == 1)

cor_therm_study1 <- cor(gay_r_study1$therm1, gay_r_study1$therm2, use = "complete.obs")

gay_r_study2 <-gay_r %>% 
  filter(study == 2)

cor_therm_study2 <- cor(gay_r_study2$therm1, gay_r_study2$therm2, use = "complete.obs")

gay_r_study1 <- gay_r_study1 %>% mutate(diff_therm2_therm1 = therm2 - therm1)
gay_r_study2 <- gay_r_study2 %>% mutate(diff_therm2_therm1 = therm2 - therm1)

sd1 <- gay_r_study1 %>% summarize(sd(diff_therm2_therm1, na.rm = TRUE))
sd2 <- gay_r_study2 %>% summarize(sd(diff_therm2_therm1, na.rm = TRUE))

cor_therm_study1
cor_therm_study2
sd1
sd2
```

talk

## Question 2.5

```{r}
study1_plot <- gay_r %>% ggplot(mapping = aes(x=therm1, y = therm2)) + geom_point(shape = 21, fill = NA) + facet_wrap(~study, labeller = labeller(study = c("1" = "Study 1", "2"
= "Study 2"))) + 
    labs(x = "Thermometer Score in Wave 1",
         y = "Thermometer Score in Wave 2",
         title = "Wave 1 & 2 Thermometer Scores, by Study 1 and Study 2") + geom_abline(slope = 1, color = "tomato")
study1_plot
```

## Question 2.6

yap

## Question 3.1

```{r}
gay_r_study2 <-gay_r %>% 
  filter(study == 2) #from question 2.4b

nrow(gay_r_study2)

sd_marriage_gay_s2w1 <- gay_r_study2 %>% 
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(sd(therm1))

mean_marriage_gay_s2w1 <- gay_r_study2 %>% 
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(mean(therm1))

sd_no_contact_s2w1 <- gay_r_study2 %>% 
  filter(treatment == "No Contact") %>%
  summarize(sd(therm1))

mean_no_contact_s2w1 <- gay_r_study2 %>% 
  filter(treatment == "No Contact") %>%
  summarize(mean(therm1))

sd_marriage_gay_s2w1
mean_marriage_gay_s2w1
sd_no_contact_s2w1
mean_no_contact_s2w1
```

## Question 3.2

```{r}


comparison_graph2 <- gay_r %>%
  ggplot(mapping = aes(x=therm1)) + 
  geom_histogram(breaks = seq(0, 100, by = 5), color ="black", fill = "blue", binwidth = 5) +
  facet_wrap(~study, labeller = labeller(study = c("1" = "Study 1", "2" = "Study 2"))) + 
    labs(x = "Thermometer Score towards Gay Men and Lesbians",
         y = "Number of Respondents",
         title = "Side-by-Side Comparison of Study 1 and Study 2 Data")
comparison_graph2
```

## Question 3.3

```{r}

```

## Question 3.4

```{r}
mean_marriage_gay_s2w2 <- gay_r_study2 %>% 
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  summarize(mean(therm2, na.rm=TRUE))

mean_no_contact_s2w2 <- gay_r_study2 %>% 
  filter(treatment == "No Contact") %>%
  summarize(mean(therm2, na.rm=TRUE))

mean_marriage_gay_s2w2
mean_no_contact_s2w2

mean_marriage_gay_s2w2 - mean_no_contact_s2w2
```

## Question 3.5

```{r}
var_gay <- gay_r_study2 %>% 
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>%
  var(therm2)

n_gay <- gay_r_study2 %>% 
  filter(treatment == "Same-Sex Marriage Script by Gay Canvasser") %>% nrow()

var_control <- gay_r_study2 %>% 
  filter(treatment == "No Contact") %>%
  var(therm2)

n_control <- gay_r_study2 %>% 
  filter(treatment == "No Contact") %>% nrow()

standard_error <- sqrt((var_gay/n_gay)+(var_control/n_control))
```

## Question 3.6

```{r}

gay_r_study2_control <- gay_r_study2 %>% filter(treatment == "No Contact")
  
  
cor_therm_study2_control12 <- cor(gay_r_study2_control$therm1, gay_r_study2_control$therm2, use = "complete.obs")

cor_therm_study2_control13 <- cor(gay_r_study2_control$therm1, gay_r_study2_control$therm3, use = "complete.obs")

cor_therm_study2_control14 <- cor(gay_r_study2_control$therm1, gay_r_study2_control$therm4, use = "complete.obs")

cor_therm_study2_control12
cor_therm_study2_control13
cor_therm_study2_control14
```

## Question 4.1

```{r}
gay_control_s1 <- gay_r %>% 
  filter(study == 1) %>% 
  filter(treatment == "No Contact")

fit_s1 <- lm(therm2 ~ therm1, data = gay_control_s1)
summary(fit_s1)
```

## Question 4.2

```{r}
gay_control_s2 <- gay_r %>% 
  filter(study == 2) %>% 
  filter(treatment == "No Contact")

fit_s2 <- lm(therm2 ~ therm1, data = gay_control_s2)
summary(fit_s2)
```

## Question 4.3

```{r}

gay_r_2 <-gay_r %>% 
  filter(study == 2) %>% 
  mutate(treated = case_when(treatment == "No Contact" ~ 0, treatment == "Same-Sex Marriage Script by Gay Canvasser" ~ 1))

fit_s3 <- lm(therm2 ~ treated, data = gay_r_2)
summary(fit_s3)

```

## Question 4.4

```{r}

fit_s4 <- lm(therm2 ~ treated + therm1, data = gay_r_2)
summary(fit_s4)
```

## Question 4.5

```{r}
predict_df1 <- data.frame(treated = 1, therm1 = 50)
predict1 <- predict(fit_s4, predict_df1)
print(predict1)

predict_df2 <- data.frame(treated = 0, therm1 = 50)
predict2 <- predict(fit_s4, predict_df2)
print(predict2)


predict1-predict2
```

The predicted difference in their Wave 2 attitudes
